---
title: "pstat-131-section-3-allester"
author: "allester"
date: '2022-10-12'
output: html_document
---

```{r}
library(tidyverse)
# install.packages('infer') # execute once then comment out

# data location
url <- 'https://raw.githubusercontent.com/pstat197/pstat197a/main/materials/labs/lab3-iteration/data/biomarker-clean.csv'

# function for outlier trimming
trim_fn <- function(x){
  x[x > 3] <- 3
  x[x < -3] <- -3
  
  return(x)
}

# read in and preprocess data
asd <- read_csv(url) %>%
  select(-ados) %>%
  # log transform
  mutate(across(.cols = -group, log10)) %>%
  # center and scale
  mutate(across(.cols = -group, ~ scale(.x)[, 1])) %>%
  # trim outliers
  mutate(across(.cols = -group, trim_fn))
```

```{r}
asd
```

# Multiple Testing with Loops

```{r}
x <- asd %>% filter(group == 'ASD') %>% pull(CHIP)
y <- asd %>% filter(group == 'TD') %>% pull(CHIP)
t.test(x, y, var.equal = F)
```

```{r}
t.test(x, y) %>% str()
```

``` {r}
n_tests <- 100
p_vals <- rep(NA, n_tests)

for(i in 1:n_tests){
  x <- asd %>% filter(group == 'ASD') %>% pull(i + 1)
  y <- asd %>% filter(group == 'TD') %>% pull(i + 1)
  p_vals[i] <- t.test(x, y, var.equal = F)$p.value
}
```

``` {r}
tibble(protein = colnames(asd)[2:(n_tests + 1)],
       p = p_vals)
```


# t-test using apply
``` {r}
# number of tests to perform
n_tests <- 100

# convert to a list
asd_list <- asd %>% 
  select(1:(n_tests + 1)) %>%
  pivot_longer(cols = -group,
               names_to = 'protein',
               values_to = 'level') %>%
  group_by(protein) %>%
  group_split()

# first entry in list
asd_list[[1]]
```

``` {r}
t.test(level ~ group, data = asd_list[[1]])
```


Make a function that perfoms a t test on the ith index of a list of dfs

```{r}
# p value for ith protein
tt_fn <- function(i){
  t.test(level ~ group, data = asd_list[[i]])$p.value
}

# check
tt_fn(1)
```

Using sapply to apply function across a vector

``` {r}
sapply(1:n_tests, tt_fn)
```

# Nesting

``` {r}
asd_nested <- asd %>%
  pivot_longer(-group, 
               names_to = 'protein', 
               values_to = 'level') %>%
  nest(data = c(level, group))

asd_nested %>% head(5)
```

# Can also use the map function to run a t test on the data frames of a tibble

``` {r}
tt_fn <- function(.df){
  t.test(level ~ group, data = .df)
}

rslt <- asd_nested %>%
  slice(1:10) %>%
  mutate(ttest.out = map(data, tt_fn))

rslt
```

Reading the t test results from the first row using the slice() function

```{r}
rslt %>% slice(1L) %>% pull(ttest.out)
```

Output ttest col as a tibble which can be unnested

``` {r}
# wrapper around infer::t_test
tt_fn <- function(.df){
  infer::t_test(.df, 
         formula = level ~ group,
         order = c('ASD', 'TD'),
         alternative = 'two-sided',
         var.equal = F)
}

# compute test results
tt_out <- asd_nested %>%
  slice(1:n_tests) %>%
  mutate(ttest = map(data, tt_fn))

# preview
tt_out %>% head(4)
```

Unnesting the tibble in theh ttest column

```{r}
tt_out %>% 
  unnest(ttest) %>%
  head(4)
```

# Adjusting p-values
``` {r}
# bonferroni correction
tt_out %>% 
  unnest(ttest) %>%
  mutate(p_adj = p_value*n_tests) %>%
  select(protein, p_value, p_adj) %>%
  arrange(p_adj) %>%
  head(4)
```

``` {r}
n_tests <- 50
p_vals <- rep(NA, n_tests)
#e_diff <- rep(NA, n_tests)

for(i in 1:n_tests){
  x <- asd %>% filter(group == 'ASD') %>% pull(i + 1)
  y <- asd %>% filter(group == 'TD') %>% pull(i + 1)
  p_vals[i] <- t.test(x, y, var.equal = F)$p.value
  e_diff[i] <- t.test(x, y, var.equal = F)$estimate
}

t.test(x, y, var.equal = F)$estimate
```

``` {r}
tibble(protein = colnames(asd)[2:(n_tests + 1)],
       p = p_vals, e = e_diff)
```


# sapply

```{r}
tt_fn <- function(i){
  t.test(level ~ group, data = asd_list[[i]])$stderr
}

# check
tt_fn(1)
```


```{r}
tt_fn <- function(i){
  test_rslt <- t.test(level ~ group, data = asd_list[[i]])
  out <- c(diff_e = as.numeric(test_rslt$estimate[1]) - as.numeric(test_rslt$estimate[2]), 
           stderr = test_rslt$stderr)
  out
}

tt_fn(1)
```

```{r}
sapply(1:50, tt_fn) %>% t() %>% as_tibble()
```


