---
title: "lab4-classification-accuracy"
author: "allester"
date: '2022-10-19'
output: pdf_document
---

```{r}
# load packages
library(tidyverse)
library(tidymodels)
library(modelr)
library(rsample)
library(yardstick)

# read data
url <- 'https://raw.githubusercontent.com/pstat197/pstat197a/main/materials/labs/lab4-logistic/data/biomarker_clean.csv'

s_star <- c("DERM", "RELT", "IgD", "PTN", "FSTL1")
biomarker <- read_csv(url) %>%
  # subset to proteins of interest and group
  select(group, any_of(s_star)) %>%
  # convert group (chr) to binary (lgl)
  mutate(class = (group == 'ASD')) %>%
  select(-group)
```

```{r}
glimpse(biomarker)
```

```{r}
# for reproducibility
set.seed(102022)

# partition data
partitions <- biomarker %>%
  initial_split(prop = 0.8)

# examine
partitions
```

```{r}
# training set
training(partitions) %>% head(4)
```

```{r}
# testing set
testing(partitions) %>% head(4)
```

```{r}
# fit glm
fit <- glm(class ~ ., 
           data = biomarker, 
           family = binomial(link = "logit"))
```

```{r}
tidy(fit)
```

# Action: Fit the model and interpret the parameter

Take IgD, it has a p-value < 0.01; thus it is statistically significant. Since it has an estimate of -0.66 it has an odds ratio of $e^{-0.66} = .51$. This means that there the IgD group has a 51%
time the odds of the non-IgD group of having ASD.

```{r}
# compute predictions on the test set
testing(partitions) %>%
  add_predictions(fit)
```

```{r}
# manually transform to probabilities
testing(partitions) %>%
  add_predictions(fit) %>%
  mutate(probs = 1/(1 + exp(-pred))) %>%
  select(class, pred, probs) %>%
  head(5)
```

```{r}
# predict on scale of response
testing(partitions) %>%
  add_predictions(fit, type = 'response') %>%
  select(class, pred) %>%
  head(5)
```

```{r}
# predict classes
testing(partitions) %>%
  add_predictions(fit, type = 'response') %>%
  mutate(pred.class = (pred > 0.5)) %>%
  select(class, pred, pred.class) %>%
  head(5)
```

```{r}
# tabulate
testing(partitions) %>%
  add_predictions(fit, type = 'response') %>%
  mutate(pred.class = (pred > 0.5)) %>%
  select(class, pred.class) %>%
  table()
```

```{r}
# store predictions as factors
pred_df <- testing(partitions) %>%
  add_predictions(fit, type = 'response') %>%
  mutate(pred.class = (pred > 0.5),
         group = factor(class, labels = c('TD', 'ASD')),
         pred.group = factor(pred.class, labels = c('TD', 'ASD'))) 

# check order of factor levels
pred_df %>% pull(group) %>% levels()
```

```{r}
# compute specificity
pred_df %>%
  specificity(truth = group, 
              estimate = pred.group,
              event_level = 'second')
```

```{r}
# sensitivity
pred_df %>%
  sensitivity(truth = group,
              estimate = pred.group,
              event_level = 'second')
```
# Action: Accuracy

```{r}
# accuracy
pred_df %>%
  accuracy(truth = group,
              estimate = pred.group,
              event_level = 'second')
```

```{r}
# define panel (arguments must be yardstick metric function names)
panel_fn <- metric_set(sensitivity, specificity)

# compute
pred_df %>%
  panel_fn(truth = group,
           estimate = pred.group,
           event_level = 'second')
```

# Action: Compute a panel of precision, recall, and F1 score.

```{r}
# define panel (arguments must be yardstick metric function names)
panel_fn <- metric_set(precision, recall, f_meas)

# compute
pred_df %>%
  panel_fn(truth = group,
           estimate = pred.group,
           event_level = 'second')
```

```{r}
pred_df %>% conf_mat(truth = group, estimate = pred.group)
```

```{r}
pred_df %>%
  roc_curve(truth = group, estimate = pred)
```

```{r}
pred_df %>%
  roc_curve(truth = group, 
            estimate = pred,
            event_level = 'second') %>%
  ggplot(aes(y = sensitivity, x = 1 - specificity)) +
  geom_path() +
  geom_abline(slope = 1, intercept = 0)
```

```{r}
# area under the curve
pred_df %>% roc_auc(truth = group, 
                    estimate = pred,
                    event_level = 'second')
```

```{r}
panel <- metric_set(roc_auc, accuracy) 

pred_df %>% panel(truth = group,
                  estimate = pred.group,
                  pred,
                  event_level = 'second')
```

